---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 144
)

options(crayon.enabled = TRUE)
knitr::knit_hooks$set(output = function(x, options) paste0(
  '<div class="sourceCode"><pre class="sourceCode"><code class="sourceCode">',
  paste0(fansi::sgr_to_html(htmltools::htmlEscape(x)), collapse = "\n"),
  "</code></pre></div>"
))
```

# measure

<!-- badges: start -->
<!-- badges: end -->

The goal of measure is to integrate automatic unit handling into data analysis workflows.

## Installation

You can install the development version of measure like so:

``` r
# install.packages("devtools")
devtools::install_github("hrryt/measure")
```

## Example

```{r libs, message=FALSE, warning=FALSE}
library(tidyverse)
library(measure)
```

Use `measure()` to make new `measure` objects and to convert between units.

Units are handled for you when carrying out arithmetic operations.

```{r dplyr}
bodyindex <- starwars |>
  mutate(
    mass = measure(mass, "kg"), # add units to a measurement
    height = measure(height, "c|m"), # vertical bar indicates SI prefix
    bmi0 = mass / height^2, # arithmetic operators preserve units
    bmi = measure(bmi0, "kg m^-2") # convert units
  ) |>
  select(name:mass, bmi0:bmi)

bodyindex |> arrange(desc(mass))
```

Units are automatically annotated onto `ggplot2` plots with 'enmeasured' scales.

Every `ggplot2` scale has a `*_measure` equivalent.

```{r ggplot2}
breaks <- scales::breaks_extended(7)

bodyindex |>
  filter(mass < measure(200, "kg")) |>
  ggplot(aes(mass, height, color = bmi)) +
  geom_point(size = 5) +
  # two methods of achieving the same thing:
  scale_x_log10(breaks = breaks) |> enmeasure_scale() +
  scale_y_log10_measure(breaks = breaks) +
  # units are appended onto the label of an 'enmeasured' scale:
  scale_color_viridis_c_measure(trans = "log10") +
  labs(color = "BMI", title = "Body Mass Index of Star Wars Characters")
```

```{r plays}
(plays <- lakers |>
  tibble() |>
  mutate(
    date = ymd(date),
    time = ms(time),
    team,
    points = measure(points, "point"),
    .keep = "used"
  ))
```

`measure` works seamlessly with `difftime` and `hms` objects, and
`lubridate` periods, intervals and durations.

Unlike most time objects, `measure` units are preserved through summary functions.

```{r time}
plays |>
  mutate(time = measure(time, "min")) |>
  group_by(date, team) |>
  summarise(
    time = max(time),
    points = sum(points),
    rate = points / time,
    .groups = "keep"
  ) |>
  group_by(team) |>
  summarise(
    time = first(time),
    points = mean(points),
    rate = median(rate)
  )
```
